Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-05-19 ThesisJP/Thesis.tex

~\\
\newpage
%\thispagestyle{empty}%
~\\%
~\\%
~\\
\noindent\large 
\textbf{Akselerering av adaptive algoritmer for ultralydavbildning ved hjelp a programmerbare grafikkprosessorer}
%\vspace{0.2cm}\\%
\\\\%
\normalsize
En rivende utvikling innen spillteknologi og tilh\o{}rende programmeringsspr\aa{}k har nylig gjort super-datamaskiner i form av et lite grafikkort (GPU) tilgjengelige for forskere. Denne utrolige \o{}kningen i beregningskraft og forbedret programmerbarhet er for \o{}yeblikket med p\aa{} \aa{} endre hvordan ultralydsystemer designes. N\aa{}r forskere utforsker nye algoritmer for ultralydavbildning er det derfor viktig at arkitekturen til parallelle akseleratorer tas h\o{}yde for. Hvis en ny, kompleks algoritme er tenkt til \aa{} kj\o{}res i sanntid, er det viktig at den kan tilpasses den programmerbare og parallelle prosesseringskjeden i moderne ultralydskannere.

@@

I siste del av avhandlingen utforsker vi hvordan GPUer kan utnyttes til \aa{} akselerere ultralydsimuleringer. Resultatet av dette arbeidet ble et simuleringsprogram hvor geometrien til en gitt ultralydprobe kan interaktivt tegnes, og hvor det resulterende trykkfeltet simuleres i sanntid.

@@

% ENGLISH ABSTRACT / SUMMARY
\CustChapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
A rapid development in computer game technology and accompanying programming languages have recently provided researchers with small personal super computers, comprised in a single graphics card (GPU). This immense rise in computational capabilities and improved programmability are currently changing how ultrasound imaging systems are designed. When researchers are exploring new algorithms for ultrasound imaging, it is therefore important to keep the architecture of parallel accelerators in mind. If a new complex algorithm is supposed to run in real time, it needs to fit the programmable and parallel pipeline of modern ultrasound scanners.

@@

Finally, we investigate how GPUs can be utilized to accelerate ultrasound simulations (\textbf{Paper\,V}). The result of this work was a simulation program where ultrasound probe geometries can be interactively drawn and where the resulting pressure field is simulated and visualized in real time.

@@

% PREFACE AND ACKNOWLEDGEMENTS
\CustChapter*{Preface}%
\addcontentsline{toc}{chapter}{Preface}
%\thispagestyle{empty}%
The present thesis is submitted in partial fulfillment of the requirements for the degree of \textit{Philosophiae Doctor} (Ph.D.) at the Faculty of Medicine at the Norwegian University of Science and Technology (NTNU). The research was funded by \textit{Medical Imaging Laboratory} (MI-Lab), and was carried out at the Department of Circulation and Medical Imaging. The main supervisor has been Professor Sverre Holm in the Department of Informatics, University of Oslo (UiO), and Professor II at MI-Lab, NTNU. Co-supervisors have been Ph.D. Stein Inge Rabben, Ph.D. Erik Steen, both at GE Vingmed Ultrasound AS, and Professor Hans Torp in the Department of Circulation and Medical Imaging, NTNU.

>>>>

~\\
\newpage
%\thispagestyle{empty}%
~\\%
~\\%
~\\
\noindent\large 
\textbf{Akselerering av adaptive algoritmer for ultralydavbildning ved hjelp av programmerbare grafikkprosessorer}
%\vspace{0.2cm}\\%
\\\\%
\normalsize
En rivende utvikling innen spillteknologi og tilh\o{}rende programmeringsspr\aa{}k har nylig gjort superdatamaskiner i form av et lite programmerbart grafikkort (GPU) tilgjengelige for forskere. Denne utrolige \o{}kningen i beregningskraft og forbedret programmerbarhet er for \o{}yeblikket med p\aa{} \aa{} endre hvordan ultralydsystemer designes. N\aa{}r forskere utforsker nye algoritmer for ultralydavbildning er det derfor viktig at arkitekturen til parallelle akseleratorer, som en GPU er et eksempel p\aa{}, tas h\o{}yde for. Hvis en ny, kompleks algoritme er tenkt til \aa{} kj\o{}res i sanntid, er det viktig at den kan tilpasses den programmerbare og parallelle prosesseringskjeden i moderne ultralydskannere.

@@

I siste del av avhandlingen utforsker vi hvordan GPUer kan utnyttes til \aa{} akselerere ultralydsimuleringer. Resultatet av dette arbeidet ble et simuleringsprogram hvor geometrien til et gitt ultralydarray kan interaktivt tegnes, og hvor det resulterende trykkfeltet simuleres i sanntid.

@@

% ENGLISH ABSTRACT / SUMMARY
\CustChapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
A rapid development in computer game technology and accompanying programming languages have recently provided researchers with small personal supercomputers, comprised in a single graphics card (GPU). This immense rise in computational capabilities and improved programmability are currently changing how ultrasound imaging systems are designed. When researchers are exploring new algorithms for ultrasound imaging, it is therefore important to keep the architecture of parallel accelerators like the GPU in mind. If a new complex algorithm is supposed to run in real time, it needs to fit the programmable and parallel pipeline of modern ultrasound scanners.

@@

Finally, we investigate how GPUs can be utilized to accelerate ultrasound simulations (\textbf{Paper\,V}). The result of this work was a simulation program where ultrasound array geometries can be interactively drawn and where the resulting pressure field is simulated and visualized in real time.

@@

% PREFACE AND ACKNOWLEDGEMENTS
\CustChapter*{Preface}%
\addcontentsline{toc}{chapter}{Preface}
%\thispagestyle{empty}%
This thesis is submitted in partial fulfillment of the requirements for the degree of \textit{Philosophiae Doctor} (Ph.D.) at the Faculty of Medicine at the Norwegian University of Science and Technology (NTNU). The research was funded by \textit{Medical Imaging Laboratory} (MI-Lab), and was carried out at the Department of Circulation and Medical Imaging. The main supervisor has been Professor Sverre Holm in the Department of Informatics, University of Oslo (UiO), and Professor II at MI-Lab, NTNU. Co-supervisors have been Ph.D. Stein Inge Rabben, Ph.D. Erik Steen, both at GE Vingmed Ultrasound AS, and Professor Hans Torp in the Department of Circulation and Medical Imaging, NTNU.
